Abstract Plots

### 2021-07-22 Abstract.Rmd
Added specificity/sensitivity analysis for dilutions. Also changed loops that go through all signatures to be more universal (get_known_signature() does not always retrieve 60 signatures).

#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:60){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}

#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:60){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:60){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:60){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

@@

>>>>

#Create list of signature matrices with assigned rownames
signature.matrix.list <- list()
for (i in 1:length(signature.names)){
  sig <- cosmic.signatures[,i]
  sig.mat <- as.matrix(sig)
  rownames(sig.mat) <- mutations
  colnames(sig.mat) <- signature.names[i]
  signature.matrix.list[[i]] <- sig.mat
}

#create vector of cosine similarities to the FFPE signature
ffpe.cosmic.cos.sims <- c()
for (i in 1:length(signature.names)){
  cos.sim <- cos_sim_matrix(signature.matrix.list[[i]], ffpe.signature)
  ffpe.cosmic.cos.sims[i] <- cos.sim[1,1]
}
```

#Scatterplot of Average AUC vs. Cosine Similarity to FFPE
```{r}
#outer for loop: loop through each COSMIC v3 signature
average.auc.values <- c()
for (x in 1:length(signature.names)){
  signature.auc.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df.1 <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df.1)
    sig.name <- class.df.colnames[3]
    classification.df.2 <- classification.df.1 %>%
      mutate(signature.indicator = ifelse(truth == sig.name, 1, 0))
    predictions <- classification.df.2[3]
    labels <- classification.df.2$signature.indicator
    pred <- prediction(predictions, labels)
    auc.perf <- performance(pred, measure = "auc")
    signature.auc.vec[y] <- auc.perf@y.values[[1]][1]
  }
  average.signature.auc <- mean(signature.auc.vec)
  average.auc.values[x] <- average.signature.auc
}
```

@@

```{r}
#outer for loop: loop through each COSMIC v3 signature
average.sensitivity.values <- c()
average.specificity.values <- c()
for (x in 1:length(signature.names)){
  signature.sensitivity.vec <- c()
  signature.specificity.vec <- c()
  #inner for loop: calculate misclassification rate 10x for each signature
  for (y in 1:10){
    #create a sample vector of 100 mutations for the COSMIC signature corresponding to number x
    sample.vec <- create_signature_sample_vector(signature.matrix.list[[x]],100)
    #Create FFPE sample vector of 100 mutations
    ffpe.sample.vec <- create_signature_sample_vector(ffpe.signature, 100)
    classification.df <- get_classification_df(list(sample.vec, ffpe.sample.vec), c(signature.names[x],"FFPE"), list(signature.matrix.list[[x]], ffpe.signature))
    class.df.colnames <- colnames(classification.df)
    sig.name <- class.df.colnames[3]
    #Count number of true positives (for sensititivity)
    sensitivity.df <- classification.df %>%
      filter(truth == sig.name & classify == sig.name)
    #Calculate sensitivity
    sensitivity.val = nrow(sensitivity.df)/100
    #Add to the vector of 10 sensitivity values
    signature.sensitivity.vec[y] = sensitivity.val
    #Count number of true negatives (for specificity)
    specificity.df <- classification.df %>%
      filter(truth == "FFPE" & classify == "FFPE")
    #Calculate specificity
    specificity.val = nrow(specificity.df)/100
    #Add to the vector of 10 specificity values
    signature.specificity.vec[y] = specificity.val
  }
  average.signature.sensitivity <- mean(signature.sensitivity.vec)
  average.sensitivity.values[x] <- average.signature.sensitivity
  average.signature.specificity <- mean(signature.specificity.vec)
  average.specificity.values[x] <- average.signature.specificity
}
```

@@

# Dilution performance characteristics

```{r}
cosmic.sig4 = signature.matrix.list[[4]]
cosmic.sig6 = signature.matrix.list[[6]]
ffpe.sig = ffpe.signature
```

## Dilution predictions

```{r}
num.mut=1000
probabilities.MMR <- vector(mode = "list", length = 99)
truth.MMR <- vector(mode = "list", length = 99)

for (x in seq(10, num.mut-1, by = 10)){
  sample.sig6 <- create_signature_sample_vector(cosmic.sig6, x)
  sig6.truth <- rep("MMR", x)
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
  ffpe.truth <- rep("FFPE", num.mut-x)
  test <- get_classification_df(list(sample.sig6, sample.ffpe), c("MMR","FFPE"),
                                list(cosmic.sig6, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "MMR",1,0))

  probabilities.MMR[[x/10]] <- classification$MMR
  truth.MMR[[x/10]] <- classification$trueBinary
}
```

```{r}
num.mut=1000
probabilities.tob <- vector(mode = "list", length = 99)
truth.tob <- vector(mode = "list", length = 99)

for (x in seq(10, num.mut-1, by = 10)){
  sample.sig4 <- create_signature_sample_vector(cosmic.sig4, x)
  sig4.truth <- rep("Tobacco", x)
  sample.ffpe <- create_signature_sample_vector(ffpe.sig, num.mut-x)
  ffpe.truth <- rep("FFPE", num.mut-x)
  test <- get_classification_df(list(sample.sig4, sample.ffpe), c("Tobacco","FFPE"),
                                list(cosmic.sig4, ffpe.sig))
  classification <- test %>%
    mutate(trueBinary = ifelse(truth == "Tobacco",1,0))

  probabilities.tob[[x/10]] <- classification$Tobacco
  truth.tob[[x/10]] <- classification$trueBinary
}
```

```{r}
test.MMR <- prediction(probabilities.MMR, truth.MMR)
test.tob <- prediction(probabilities.tob, truth.tob)
```

## Sensitivity

### Sensitivity/MMR

```{r}
AVG.sens.MMR <- vector(mode = "list", length = 99)

for (iteration in 1:20){
  sens.MMR <- performance(test.MMR, measure = "sens")

  for(ind in 1:length(sens.MMR@x.values)){
    thresh = as.vector(slot(sens.MMR, "x.values")[[ind]])
    sens = as.vector(slot(sens.MMR, "y.values")[[ind]])
    if (length(thresh) > 2){
      if (identical(which(thresh<=0.5), integer(0))){
        AVG.sens.MMR[[ind]][iteration] = NA
      }
      else{
        low.ind = max(which(thresh >=0.5))
        high.ind = min(which(thresh <=0.5))
        #point slop formula of a line
        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
        AVG.sens.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
      }
    }
    else{
      AVG.sens.MMR[[ind]][iteration] = NA
    }
  }
}

sens.MMR.vec = c()
for (avg in 1:length(AVG.sens.MMR)){
  sens.MMR.vec[avg] = mean(AVG.sens.MMR[[avg]], na.rm = TRUE)
}
```

### Sensitivity/Tobacco

```{r}
AVG.sens.tob <- vector(mode = "list", length = 99)

for(iteration in 1:20){
  sens.tob <- performance(test.tob, measure = "sens")

  for(ind in 1:length(sens.tob@x.values)){
    thresh = as.vector(slot(sens.tob, "x.values")[[ind]])
    sens = as.vector(slot(sens.tob, "y.values")[[ind]])
    if (length(thresh) > 2){
      if (identical(which(thresh<=0.5), integer(0))){
        AVG.sens.tob[[ind]][iteration] = NA
      }
      else{
        low.ind = max(which(thresh >=0.5))
        high.ind = min(which(thresh <=0.5))
        #point slop formula of a line
        slope = (sens[high.ind] - sens[low.ind]) / (thresh[high.ind] - thresh[low.ind])
        AVG.sens.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + sens[low.ind]
      }
    }
    else{
      AVG.sens.tob[[ind]][iteration] = NA
    }
  }
}

sens.tob.vec = c()
for (avg in 1:length(AVG.sens.tob)){
  sens.tob.vec[avg] = mean(AVG.sens.tob[[avg]], na.rm = TRUE)
}
```

### Sensitivity across signatures

```{r}
plot(1:99, sens.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
     ylab = "Sensitivity value", type = "l", main = "Comparing performance characteristics")
lines(1:99, sens.tob.vec, lty = 2)
abline(h = 0.85, col = "red")
legend("bottomright", c("MMR sensitivity", "Tobacco sensitivity", 
                        "Satisfactory Sensitivity"),
       lty = c(1,2,1), col = c("black", "black", "red"))
```

## Specificity

### Specificity/MMR

```{r}
AVG.spec.MMR <- vector(mode = "list", length = 99)

for(iteration in 1:20){
  spec.MMR <- performance(test.MMR, measure = "spec")

  for(ind in 1:length(spec.MMR@x.values)){
    thresh = as.vector(slot(spec.MMR, "x.values")[[ind]])
    spec = as.vector(slot(spec.MMR, "y.values")[[ind]])
    if (length(thresh) > 2){
      if (identical(which(thresh<=0.5), integer(0))){
        AVG.spec.MMR[[ind]][iteration] = NA
      }
      else{
        low.ind = max(which(thresh >=0.5))
        high.ind = min(which(thresh <=0.5))
        #point slop formula of a line
        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
        AVG.spec.MMR[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
      }
    }
    else{
      AVG.spec.MMR[[ind]][iteration] = NA
    }
  }
}

spec.MMR.vec = c()
for (avg in 1:length(AVG.spec.MMR)){
  spec.MMR.vec[avg] = mean(AVG.spec.MMR[[avg]], na.rm = TRUE)
}
```

### Specificity/Tobacco

```{r}
AVG.spec.tob <- vector(mode = "list", length = 99)

for (iteration in 1:20){
  spec.tob <- performance(test.tob, measure = "spec")

  for(ind in 1:length(spec.tob@x.values)){
    thresh = as.vector(slot(spec.tob, "x.values")[[ind]])
    spec = as.vector(slot(spec.tob, "y.values")[[ind]])
    if (length(thresh) > 2){
      if (identical(which(thresh<=0.5), integer(0))){
        AVG.spec.tob[[ind]][iteration] = NA
      }
      else{
        low.ind = max(which(thresh >=0.5))
        high.ind = min(which(thresh <=0.5))
        #point slop formula of a line
        slope = (spec[high.ind] - spec[low.ind]) / (thresh[high.ind] - thresh[low.ind])
        AVG.spec.tob[[ind]][iteration] = slope*(0.5 - thresh[low.ind]) + spec[low.ind]
      }
    }
    else{
      AVG.spec.tob[[ind]][iteration] = NA
    }
  }
}

spec.tob.vec = c()
for (avg in 1:length(AVG.spec.tob)){
  spec.tob.vec[avg] = mean(AVG.spec.tob[[avg]], na.rm = TRUE)
}
```

### Specificity across all signatures

```{r}
plot(1:99, spec.MMR.vec, ylim = c(0,1), xlab = "Percent of each signature", 
     ylab = "Specificity value", type = "l", main = "Comparing performance characteristics")
lines(1:99, spec.tob.vec, lty = 2)
abline(h = 0.85, col = "red")
legend("bottomleft", c("MMR specificity", "Tobacco specificity", 
                       "Satisfactory Specificity"),
       lty = c(1,2,1), col = c("black", "black", "red"))
```
